Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-18 BN_discretizer_paper.tex
woo2

The following four principles for the optimal discretization policy, motivated by MODL, enable the formulation of $P(M)$ and $P(D \mid M)$:
\begin{enumerate}
\item A discretization edge located at $(x_i + x_{i+1})/2$ has probability
\begin{equation}
1 - {\exp(- L \cdot {{x_{i+1} - x_i}\over{x_N - x_1}})},
\end{equation}
where $L$ is the largest cardinality number of discrete variables in $X$'s Markov blanket. \tim{Where does this come from? Is there an intuitive explanation?}\mykel{$\bigstar$}
\item For a given interval of $M$, every distribution of $\Pi_X$ is equiprobable.
\item For each pair of $(C_i,\boldsymbol{S_i})$ and a given interval of $M$, every distribution of $C_i$ with a value of $\boldsymbol{S_i}$ is equiprobable.\mykel{$\bigstar$}
\item The distributions of each $C_i$ and $\Pi_X$ in each interval of $M$ with each value of $\boldsymbol{S_i}$ are independent from each other.\mykel{$\bigstar$}
\end{enumerate}

From the first principle one obtains
\begin{equation}
  \label{eq:p_M}
  P(M) = \prod_{i=1}^k \curly{ \brock{1 - {\exp(- L \cdot {{x_{s_i+1} - x_{s_i}}\over{x_N - x_1}})}} \cdot \exp\paren{L \cdot {{x_{s_{i}} - x_{s_{i-1} + 1}}\over{x_N - x_1}}}},
\end{equation}

\noindent
where $s_i = \sum_{j=1}^k n_j$ and $s_k = N$, $s_0 = 0$.
The Bayesian network graph structure causes the likelihood term $P(D \mid M)$ to factorizes according to $X$'s Markov blanket:

\begin{equation}
  \label{eq:p_D_given_M}
  P\paren{D \mid M} \propto P\paren{D_{\Pi_x} \mid M} \cdot \prod_{i = 1}^{n_c} P\paren{D_{C_i} \mid M, D_{\boldsymbol{S_i}}}\text{.}
\end{equation}

For example, in Figure~\ref{fig:example_factorization}, the corresponding $P(D \mid M)$ is factorized according to
\begin{equation}
P\paren{D_{ P_1,P_2,P_3 } \mid M} \cdot P\paren{ D_{ C_1 } \mid M,D_{\boldsymbol{S_1}}} \cdot P\paren{D_{C_2} \mid M,D_{ S_2  }}\text{.}
\end{equation}

\begin{figure}[ht]
\centering
  \label{fig:example_factorization}
  %\begin{tabular}{cc}
    \input{graph1}
  %\end{tabular}
  \caption{Factorization of $P(D \mid M)$}
\end{figure}

The concept behind the factorization also the motivation behind forward sampling in a Bayesian network.
The parents of $X$ are independent of the children given $X$.
The parents are not necessarily individually independent, and thus the parental term $P(D_{\Pi_x} \mid M)$ cannot be factored further.
The children of $X$ are similarly independent given $X$ and its spouses $\boldsymbol{S_i}$, leading to their factored product $\prod_{i = 1}^{n_c} P(D_{C_i} \mid M, D_{\boldsymbol{S_i}})$.
Each component in the decomposition be evaluated based on a discretization policy $M = [n_1,n_2,\ldots,n_k]$ and dataset $D$.

\subsubsection{Evaluation of $P(D_{\Pi_x} \mid M)$}
Let $J_P = \prod_{i=1}^{n_p} \| P_i \|$. It follows that

\begin{equation}
\label{eq:likelihood_one}
P(D_{\Pi_X} \mid M) = \prod_{i=1}^k  {{1}\over{{n_i + J_P - 1}\choose{J_P -1}}}
{{1}\over{ {{{n}_i}!}\over{ {n^{(p)}_{i,1} !} {n^{(p)}_{i,2} !} \cdots {n^{(p)}_{i,J_P} !}}  }}\text{,}
\end{equation}

@@

\begin{algorithm}
\caption{Discretization of one continuous variable in a Bayesian network}
\label{alg:disc_one}
\begin{algorithmic}[5]
\Function{Discretize}{$D$, $G$}
\State $N \leftarrow$ the number of instances
\State $N' \leftarrow$ the number of unique values in $D_X$
\State $H \leftarrow$ an $N \times N$ matrix such that $H[u,v] = h(u,v)$ per Algorithm~\ref{alg:h} in the appendix
\State $b \leftarrow$ the increasing sequence of $M$ integers as previously defined
\State $L \leftarrow$ the largest cardinality over all discrete variables in the Markov blanket of $X$
\State $S[u] \leftarrow$ the optimal objective value of subproblems with instances from $1$ to $u$
\State $M[u] \leftarrow$ the optimal discretization for subproblems with instances from $1$ to $u$
\State $W[u]  \leftarrow - \log\brock{1 - {\exp\paren{- L \cdot{ {{x_{b(u)+1} - x_{b(u)}}\over{x_N - x_1}}}}}}$ for $u = 1,2, \ldots,N'-1$ and $L(N') \leftarrow 0$
\For {$v \leftarrow 1$ to $N'$}
\If {$v = 1$}
\State $S[v] \leftarrow g \left(1,b[v] \right) + L[v]$
\State $M[v] \leftarrow \{ ({x_{b[v]} + x_{b[v]+1}}) / 2\}$
\Else
\State $s \leftarrow \infty$ and $boundary \leftarrow \infty$
\For {$u \leftarrow 1$ to $v$}
\State $s' \leftarrow S[u] + g \left( b[u]+1,b[v] \right) +  {L \cdot {{x_{b[v]} - x_{b[u] + 1}}\over{x_N - x_1}}} + W[v]$
\If {$s' < s$}
\State $s \leftarrow s'$
\State $boundary \leftarrow ({x_{b[u]} + x_{b[u]+1}}) / 2$
\EndIf
\EndFor
\State $S[v] \leftarrow s$
\State $M[v] \leftarrow M[u] \cup \{ boundary\}$
\EndIf
\EndFor
 \State \Return $M$
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
  \caption{Discretization of multiple continuous variables \todo{Fix label reference number}}
  \label{alg:disc_two}
  \begin{algorithmic}[5]
  \Function{Discretize}{$D''$, $G$, $C$, $u_{cycle}$}
    \State
    \State $M[i] \leftarrow$ the discretization policy for the $i$th variable
    \State $n \leftarrow$ the number of variables in the Bayesian network
    \State $D^*_i \leftarrow M[i] (D_i)$, list of samples $D_i$ discretized according to $M[i]$
    \State $D^* \leftarrow $ the discretized data for all variables, where $D^*_i \leftarrow D''_i$ if $i \notin C$
    \State $k \leftarrow median\{ |v|, v\notin C\}$
    \State
    \For {$i = 1$ to $n$}
      \If {$i \in C$}
        \State $M[i] \leftarrow$  equal-width discretization with $k$ intervals
        \State $D^*_i \leftarrow  M[i] (D''_i)$
      \EndIf
    \EndFor
    \State
    \State $cycle \leftarrow 0$
    \While {$M$ is not converged  \textbf{and} $cycle \leq u_{cycle}$}
      \State {$cycle  \leftarrow cycle  + 1$}
      \For {$j = 1$ to $|C|$}
        \State $D^*_{\backslash C(j)} \leftarrow D^*$ without $D^*_{C(j)}$.
        \State {$M[C(j)] \leftarrow $ DISCRETIZE($D_{C(j)}$,$D^*_{ \backslash C(j)}$,graph)}
        \State {$D^*_{C(j)}  \leftarrow$ $M[C(j)](D''_{C(j)})$}
      \EndFor
    \EndWhile
    \State \Return $M$
  \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
  \caption{ Learning a discrete-valued Bayesian network}
  \label{alg:structure_learn}
  \begin{algorithmic}[5]
  \Function{Learn\_DBN}{$D''$, $C$, $order$, $u_{parent}$, $C$, $u_{cycle}$}
    \State $n \leftarrow$ the number of variables in the Bayesian network
    \State $k \leftarrow median\{ \|v\|, v\notin C\}$
    \State $M[i] \leftarrow$ the discretization policy for the $i$th variable
    \State $D^*_i \leftarrow M[i] (D_i)$, list of samples $D_i$ discretized according to $M[i]$
    \State $D^* \leftarrow $ the discretized data for all variables, where $D^*_i \leftarrow D''_i$ if $i \notin C$
    \State $G \leftarrow$ the initial edgeless graph structure
    \For {$i = 1$ to $n$}
      \If {$i \in C$}
        \State $M[i] \leftarrow$  equal-width discretization with $k$ intervals
        \State $D^*_i \leftarrow  M[i] (D''_i)$
      \EndIf
    \EndFor
    \For {$i =1$ to $n$}
      \State $P_{old} \leftarrow f(X^*_i,\Pi_{X^*_i})$: Equation~\ref{eq:f_x_p}
      \State OKToProceed $\leftarrow$ \textbf{true}
      \While {OKToProceed \textbf{and} $\|\Pi_{X^*_i}\| < u_{parent}$}
        \State $Y \leftarrow$ an element from the set $order[1:i] \backslash \Pi_X$
        \State $P_{new} \leftarrow f(X^*_i,\Pi_{X^*_i} \cup Y)$
        \If {$P_{new} > P_{old}$}
          \State $P_{old} \leftarrow P_{new}$
          \State $\Pi_{X_i} \leftarrow \Pi_{X_i} \cup Y $
          \State $M \leftarrow$ {DISCRETIZE}({$D''$, $G$, $C$, $u_{cycle}$}): Algorithm~\ref{alg:disc_two}
          \State $D* = M(D'')$
        \Else
          \State OKToProceed $\leftarrow$ \textbf{false}
        \EndIf
      \EndWhile
    \EndFor
    \State \Return $G$, $M$
  \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\begin{equation}
\label{eqn:MDL}
\begin{aligned}
\frac{1}{2} \log(N) \left\lbrace  \lvert \Pi_{X_i} \rvert (\lvert X^*_i \rvert - 1) +
 {\sum_{j,X_i \in \Pi_{X_i}}} \lvert \Pi_{X^*_j} \rvert (\lvert X_j \rvert - 1) \right\rbrace + \log(\lvert {X^*_i} \rvert) \\
 + (N_i-1) H \left( {{|X^*_i| - 1}\over{N_i -1}}  \right) -N \cdot \left[ I(X^*_i,\Pi_{X_i}) + {\sum_{j,X_i \in \Pi_{X_i}}} I(X_j, \Pi_{X^*_j}) \right]\text{,}
\end{aligned}
\end{equation}

where $I(A,B)$ is the mutual information between the discrete variables $A$ and $B$, ${H(p) = -p \log(p) - (1-p) \log(1-p)}$ \tim{does $H(p)$ have a name? Entropy?}, and \todo{$X^*_i$ is the discretized version of the $i$th sample of $X$}.
The optimal discretization policy is found with dynamic programming in a similar manner as with the Bayesian method.

@@

\begin{figure}[ht]
\centering
      \input{graph_auto_exp1_distr_1_3}
  \caption{Comparison of the raw data and discretization result on variable $A$ and $C$ by the optimal Bayesian approach. The marginal probabilities of discretized areas match the distribution of the taw data very well.}
  \label{fig:auto_exp1_distr_1_3}
\end{figure}

@@

\begin{figure}[ht]
\centering
      \input{graph_auto_exp2_distr_1_3}
  \caption{Comparison of the raw data and discretization result on variable $A$ and $C$ in Figure \ref{fig:auto_graph_2}. Although the number of discretization edges is less than in Figure \ref{fig:auto_exp1_distr_1_3}, it still matches the raw data distribution very well.}
  \label{fig:auto_exp2_distr_1_3}
\end{figure}

@@

The Bayesian network in Figure \ref{fig:wine_graph_1} is obtained by equal-width discretizing each continuous variable into 3 intervals and running K2 a thousand times. Table \ref{table:wine_disc_table_1} shows a comparison between the optimal Bayesian approach and the MDL principle discretization method.
The discretization result obtained by the optimal Bayesian approach is better than the MDL discretization method in terms of likelihood. The heuristic method discretizes variable $G$, $H$ and $J$ into one less interval than the optimal Bayesian approach, and has one more number of intervals on variable $M$.
Although it returns different numbers of intervals and different positions of discretization edges with the optimal Bayesian approach, it still has a good discretization result in terms of likelihood. Furthermore, if running time is the main concern, the heuristic method can be considered a better approach in this case.

@@

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{
      \input{graph_wine_1}}
  \caption{Structure from prediscretized K2 structure learning to compare the methods for Wine dataset}
  \label{fig:wine_graph_1}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  %\begin{tabular}{cc}
  \scalebox{0.8}{
   \input{graph_housing_1}}
   %\end{tabular}
  \caption{Structure from prediscretized K2 structure learning to compare the methods for Housing dataset}
  \label{fig:housing_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{Discretization result of Housing dataset based on the graph in Figure~\ref{fig:housing_graph_1}. The first twelve rows show the numbers of intervals after discretization by the three method. The thirteenth row is the log-likelihood of the dataset based on the discretizatio results. The heuristic method is over sensitive and discretizes some variables into too many intervals. The MDL method is less sensitive and discretizes most variables by only one interval. The optimal Bayesian approach has a good control on number of intervals. \tim{Fig. vs. Figure}}
  \scalebox{1.0}{
  \input{table_housing_1.tex}
  }

@@

\begin{figure}[ht]
\centering
    % \begin{tabular}{cc}
      \input{graph_housing_exp1_distr_3_5}
    % \end{tabular}
  \caption{The discretization result on variable $C$ and $E$ of the network in Figure~\ref{fig:housing_graph_1}. The scatter points are jittered due to the repeated values in the dataset. Furthermore, these repeated values lead to high number of discretization edges, since repeated values can easily form discretized areas by their own.}
  \label{fig:housing_exp1_distr_3_5}
\end{figure}

\begin{figure}[ht]
\centering
     %\begin{tabular}{cc}
    \input{graph_housing_exp1_distr_8_5}
     %\end{tabular}
  \caption{The discretization result on variable $H$ and $E$ of the network in Figure~\ref{fig:housing_graph_1}. Data points are not over repeated on $H$, thus the number of edges on $H$ remains desirable.}
  \label{fig:housing_exp1_distr_8_5}
\end{figure}

@@

\begin{figure}[ht]
    % \begin{tabular}{cc}
    \centering
    \scalebox{0.8}{
      \input{graph_housing_2}}
    % \end{tabular}
  \caption{The learned discrete-valued Bayesian network from the Housing dataset. Variables $C$, $E$, $J$, $K$ have high numbers of intervals after discretization, since they are in a Markov blanket.
  \tim{Var with 1 interval?}
  }
  \label{fig:housing_graph_2}
\end{figure}

@@

\begin{figure}[ht]
\centering
     %\begin{tabular}{cc}
      \input{graph_housing_exp2_distr_3_5}
     %\end{tabular}
  \caption{The discretization result on variable $C$ and $E$ of the network in Figure~\ref{fig:housing_graph_2}. The high number of discretization edges is due to the repeated values on the dataset in variable $C$ and $E$.}
  \label{fig:housing_exp2_distr_3_5}
\end{figure}

@@

To quantatively discuss the sensitivity of the three methods, consider the Bayesian network in Figure \ref{fig:exp_discuss}, where $X$ is continuous-valued and $P_1$ and $P_2$ is discrete-valued. Let $\Pi_X = \{  P_1, P_2 \}$. Assume $D_X = \{1.0,2.0,3.0, \ldots N \}$. Therefore, the corresponding objective functions to discretize $X$ in the three methods are:
{\begin{small}
\begin{equation}
\begin{aligned}
\text{Obj}_{\text{MDL}} & = \overbrace{Z_1 \cdot k + \log(k) + \log {{N +k - 1}\choose{k -1}} }^{\text{Penalty Term}}+   \overbrace{\log(N) \cdot I(X^*, \Pi_X)}^{\text{Disc. Edge Positions Term}} \\
\text{Obj}_{\text{Bayesian}} &= \overbrace{Z_2 \cdot k + \log {{n_i + J_P - 1}\choose{J_P -1}}}^{\text{Penalty Term}} + \overbrace{\sum^k_{i=1} \log \left(   {{n_i !}\over{n^{(P)}_{i,1}! n^{(P)}_{i,2}! \cdots n^{(P)}_{i,J_P}!    }} \right) }^{\text{Disc. Edge Positions Term}}\\
\text{Obj}_{\text{Heuristic}} &= \overbrace{Z_2 \cdot k + \log {{n_i + J_{P_1} - 1}\choose{J_{P_1} -1}} + \log {{n_i + J_{P_2} - 1}\choose{J_{P_2} -1}} }^{\text{Penalty Term}}+\\
& \overbrace{\sum^k_{i=1}  \left[ \log \left(   {{n_i !}\over{n^{(P_1)}_{i,1}! n^{(P_1)}_{i,2}! \cdots n^{(P_1)}_{i,J_{P_1}}    }} \right)  + \log \left(   {{n_i !}\over{n^{(P_2)}_{i,1}! n^{(P_2)}_{i,2} \cdots n^{(P_2)}_{i,J_{P_2}} }} \right) \right]}^{\text{Disc. Edge Positions Term}}
\end{aligned}
\end{equation}
\end{small}}
, where $Z_1$ and $Z_2$ are two constants, $k$ is the number of intervals after discretization, $X^*$ is the discretized version of $X$, and $I(X^*, \Pi_X)$ is the mutual information between $X^*$ and its parent variables. Note that the third term in $\text{Obj}_{\text{MDL}}$ is approximated by $H(p)$ as in Equation \ref{eqn:MDL} in the original paper \citep{Friedman_1996}.

>>>>

Let $D_{\bm{Y}}$ be the subset of the data corresponding to $\bm{Y}$.
Four principles for the optimal discretization policy enable the formulation of $P(\Lambda)$ and $P(D_{-\cX} \mid \Lambda)$, where the probability of the data associated with the target variable $P(D_{\cX})$ is already captured in $P(\Lambda)$.
The four principles, motivated by MODL \todo{and Lustgarden}, are:

\begin{enumerate}
\item The prior probability of a discretization edge between two consecutive unique values $\cu_i$ and $\cu_{i+1}$ is proportional to their difference:

  \begin{equation}
  1 - {\exp(- L \cdot {\frac{\cu_{i+1} - \cu_i}{\cu_m - \cu_1}})}\text{,}
  \end{equation}

where $L$ is the largest number of intervals among discrete variables in $\cX$'s Markov blanket.
This encourages edges between well-separated values over edges between closely packed samples.

\item For a given discretization interval, every distribution over $\Pa_{\cX}$ is equiprobable.
\item For each pair $\anglebrackets{C_i,\bm{S}_i}$ and a given discretization interval, every distribution over $C_i$ given an instance of $\bm{S}_i$ is equiprobable.
\item The distributions over each child and each parent in each discretization interval with each instance of $\bm{S}_i$ are independent from each other.
\end{enumerate}

From the first principle one obtains
\begin{equation}
  \label{eq:p_M}
  P(\Lambda) = \prod_{i=1}^k
  \curly{
    \brock{
      1 - \exp\paren{
        - L \cdot \frac{
                         \cu_{i+1} - \cu_i
                       }{
                         \cu_m - \cu_1
                        }
      }
    } \cdot \exp\paren{
      L \cdot \frac{\cx_{s_{i}} - \cx_{s_{i-1} + 1}}{\cx_n - \cx_1}}
    }\text{.}
\end{equation}

The Bayesian network graph structure causes the likelihood term $P(D \mid \Lambda)$ to factorizes according to $\cX$'s Markov blanket:

\begin{equation}
  \label{eq:p_D_given_M}
  P\paren{D_{-\cX} \mid \Lambda} \propto P\paren{D_{\Pa_\cX} \mid \Lambda} \cdot \prod_{C_i} P\paren{D_{C_i} \mid \Lambda, D_{\bm{S}_i}}\text{.}
\end{equation}

For example, in Figure~\ref{fig:example_factorization}, $P(D_{-\cX} \mid \Lambda)$ factors according to

\begin{equation}
  P\paren{D_{P_1,P_2,P_3} \mid \Lambda} \cdot P\paren{ D_{ C_1 } \mid \Lambda, D_{\bm{S}_1} } \cdot P\paren{D_{C_2} \mid \Lambda, D_{\bm{S}_2}}\text{.}
\end{equation}

\begin{figure}[ht]
  \centering
  \begin{tabular}{cc}
    \input{graph1}
  \end{tabular}
  \caption{Factorization of $P(D \mid \Lambda)$}
  \label{fig:example_factorization}
\end{figure}

The concept behind the factorization is also the motivation behind forward sampling in a Bayesian network.
The parents of $\cX$ are independent of the children given $\cX$.
The parents are not necessarily individually independent, and thus the parental term $P(D_{\Pa_{\cX}} \mid \Lambda)$ cannot be factored further.
The children of $\cX$ are similarly independent given $\cX$ and the corresponding spouses, leading to their factored product $\prod_{C_i} P(D_{C_i} \mid \Lambda, D_{\bm{S}_i})$.
Each component in the decomposition can be evaluated given a discretization policy and a dataset.

\subsubsection{Evaluation of $P\paren{D_{\Pa_{\cX}} \mid \Lambda}$}

Let $J_P = \prod_{i=1}^{n_p} \| P_i \|$. It follows that

\begin{equation}
  \label{eq:likelihood_one}
  P\paren{D_{\Pa_{\cX}} \mid \Lambda} = \prod_{i=1}^k
    \frac{1}{{{n_i + J_P - 1}\choose{J_P - 1}}}
    \frac{1}{
      \frac{
        {n_i}!
      }{
        {n^{(p)}_{i,1}}! \; {n^{(p)}_{i,2}}! \; \cdots \; {n^{(p)}_{i,J_P}}!
      }
    }\text{,}
\end{equation}

@@

\begin{algorithm}
\caption{Discretization of one continuous variable in a Bayesian network}
\label{alg:disc_one}
\begin{algorithmic}[5]
\Function{Discretize}{$D$, $G$}
\State
\State $N \leftarrow$ the number of instances
\State $N' \leftarrow$ the number of unique values in $D_X$
\State $H \leftarrow$ an $N \times N$ matrix such that $H[u,v] = h(u,v)$ per Algorithm~\ref{alg:h} in the appendix
\State $b \leftarrow$ the increasing sequence of $\Lambda$ integers as previously defined
\State $L \leftarrow$ the largest cardinality over all discrete variables in the Markov blanket of $X$
\State $S[u] \leftarrow$ the optimal objective value of subproblems with instances from $1$ to $u$
\State $M[u] \leftarrow$ the optimal discretization for subproblems with instances from $1$ to $u$
\State $W[u]  \leftarrow - \log\brock{1 - {\exp\paren{- L \cdot{ {{x_{b(u)+1} - x_{b(u)}}\over{x_N - x_1}}}}}}$ for $u = 1,2, \ldots,N'-1$ and $L(N') \leftarrow 0$
\State
\For {$v = 1$ to $N'$}
\If {$v = 1$}
\State $S[v] \leftarrow g \left(1,b[v] \right) + L[v]$
\State $M[v] \leftarrow \{ ({x_{b[v]} + x_{b[v]+1}}) / 2\}$
\Else
\State $s \leftarrow \infty$ and $boundary \leftarrow \infty$
\For {$u = 1$ to $v$}
\State $s' \leftarrow S[u] + g \left( b[u]+1,b[v] \right) +  {L \cdot {{x_{b[v]} - x_{b[u] + 1}}\over{x_N - x_1}}} + W[v]$
\If {$s' < s$}
\State $s \leftarrow s'$
\State $boundary \leftarrow ({x_{b[u]} + x_{b[u]+1}}) / 2$
\EndIf
\EndFor
\State $S[v] \leftarrow s$
\State $M[v] \leftarrow M[u] \cup \{ boundary\}$
\EndIf
\EndFor
 \State \Return $\Lambda$
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
  \caption{Discretization of multiple continuous variables \todo{Fix label reference number}}
  \label{alg:disc_two}
  \begin{algorithmic}[5]
  \Function{Discretize}{$D''$, $G$, $C$, $u_{cycle}$}
    \State
    \State $M[i] \leftarrow$ the discretization policy for the $i$th variable
    \State $n \leftarrow$ the number of variables in the Bayesian network
    \State $D^*_i \leftarrow M[i] (D_i)$, list of samples $D_i$ discretized according to $M[i]$
    \State $D^* \leftarrow $ the discretized data for all variables, where $D^*_i \leftarrow D''_i$ if $i \notin C$
    \State $k \leftarrow median\{ |v|, v\notin C\}$
    \State
    \For {$i = 1$ to $n$}
      \If {$i \in C$}
        \State $M[i] \leftarrow$  equal-width discretization with $k$ intervals
        \State $D^*_i \leftarrow  M[i] (D''_i)$
      \EndIf
    \EndFor
    \State
    \State $cycle \leftarrow 0$
    \While {$\Lambda$ is not converged  \textbf{and} $cycle \leq u_{cycle}$}
      \State {$cycle  \leftarrow cycle  + 1$}
      \For {$j = 1$ to $|C|$}
        \State $D^*_{\backslash C(j)} \leftarrow D^*$ without $D^*_{C(j)}$.
        \State {$M[C(j)] \leftarrow $ DISCRETIZE($D_{C(j)}$,$D^*_{ \backslash C(j)}$,graph)}
        \State {$D^*_{C(j)}  \leftarrow$ $M[C(j)](D''_{C(j)})$}
      \EndFor
    \EndWhile
    \State \Return $\Lambda$
  \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
  \caption{ Learning a discrete-valued Bayesian network}
  \label{alg:structure_learn}
  \begin{algorithmic}[5]
  \Function{Learn\_DBN}{$D''$, $C$, $order$, $u_{parent}$, $C$, $u_{cycle}$}
    \State $n \leftarrow$ the number of variables in the Bayesian network
    \State $k \leftarrow median\{ \|v\|, v\notin C\}$
    \State $M[i] \leftarrow$ the discretization policy for the $i$th variable
    \State $D^*_i \leftarrow M[i] (D_i)$, list of samples $D_i$ discretized according to $M[i]$
    \State $D^* \leftarrow $ the discretized data for all variables, where $D^*_i \leftarrow D''_i$ if $i \notin C$
    \State $G \leftarrow$ the initial edgeless graph structure
    \For {$i = 1$ to $n$}
      \If {$i \in C$}
        \State $M[i] \leftarrow$  equal-width discretization with $k$ intervals
        \State $D^*_i \leftarrow  M[i] (D''_i)$
      \EndIf
    \EndFor
    \For {$i =1$ to $n$}
      \State $P_{old} \leftarrow f(X^*_i,\Pi_{X^*_i})$: Equation~\ref{eq:f_x_p}
      \State OKToProceed $\leftarrow$ \textbf{true}
      \While {OKToProceed \textbf{and} $\|\Pi_{X^*_i}\| < u_{parent}$}
        \State $Y \leftarrow$ an element from the set $order[1:i] \backslash \Pi_X$
        \State $P_{new} \leftarrow f(X^*_i,\Pi_{X^*_i} \cup Y)$
        \If {$P_{new} > P_{old}$}
          \State $P_{old} \leftarrow P_{new}$
          \State $\Pi_{X_i} \leftarrow \Pi_{X_i} \cup Y $
          \State $M \leftarrow$ {DISCRETIZE}({$D''$, $G$, $C$, $u_{cycle}$}): Algorithm~\ref{alg:disc_two}
          \State $D* = M(D'')$
        \Else
          \State OKToProceed $\leftarrow$ \textbf{false}
        \EndIf
      \EndWhile
    \EndFor
    \State \Return $G$, $\Lambda$
  \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\begin{equation}
\begin{aligned}
\frac{1}{2} \log(N) \left\lbrace  \lvert \Pi_{X_i} \rvert (\lvert X^*_i \rvert - 1) +
 {\sum_{j,X_i \in \Pi_{X_i}}} \lvert \Pi_{X^*_j} \rvert (\lvert X_j \rvert - 1) \right\rbrace + \log(\lvert {X^*_i} \rvert) \\
 + (N_i-1) H \left( {{|X^*_i| - 1}\over{N_i -1}}  \right) -N \cdot \left[ I(X^*_i,\Pi_{X_i}) + {\sum_{j,X_i \in \Pi_{X_i}}} I(X_j, \Pi_{X^*_j}) \right]\text{,}
\end{aligned}
\end{equation}

where $I(A,B)$ is the mutual information between the discrete variables $A$ and $B$, ${H(p) = -p \log(p) - (1-p) \log(1-p)}$, and \todo{$X^*_i$ is the discretized version of the $i$th sample of $X$}.
The optimal discretization policy is found with dynamic programming in a similar manner as with the Bayesian method.

@@

\begin{figure}[ht]
      \input{graph_auto_exp1_distr_1_3}
  \caption{Comparison of the raw data and discretization result on variable $A$ and $C$ by the optimal Bayesian approach. The marginal probabilities of discretized areas match the distribution of the taw data very well.}
  \label{fig:auto_exp1_distr_1_3}
\end{figure}

@@

\begin{figure}[ht]
      \input{graph_auto_exp2_distr_1_3}
  \caption{Comparison of the raw data and discretization result on variable $A$ and $C$ in Figure \ref{fig:auto_graph_2}. Although the number of discretization edges is less than in Figure \ref{fig:auto_exp1_distr_1_3}, it still matches the raw data distribution very well.}
  \label{fig:auto_exp2_distr_1_3}
\end{figure}

@@

The Bayesian network in Figure \ref{fig:wine_graph_1} is obtained by equal-width discretizing each continuous variable into 3 intervals and running K2 a thousand times. Table \ref{table:wine_disc_table_1} shows a comparison between the optimal Bayesian approach and the MDL principle discretization method.
The discretization result obtained by the optimal Bayesian approach is better than the MDL discretization method in terms of likelihood. The heuristic method discretizes variable $G$, $H$ and $J$ into one less interval than the optimal Bayesian approach, and has one more number of intervals on variable $\Lambda$.
Although it returns different numbers of intervals and different positions of discretization edges with the optimal Bayesian approach, it still has a good discretization result in terms of likelihood. Furthermore, if running time is the main concern, the heuristic method can be considered a better approach in this case.

@@

\begin{figure}[ht]
  \centering
      \input{graph_wine_1}
  \caption{Structure from prediscretized K2 structure learning to compare the methods for Wine dataset}
  \label{fig:wine_graph_1}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \begin{tabular}{cc}
   \input{graph_housing_1}
   \end{tabular}
  \caption{Structure from prediscretized K2 structure learning to compare the methods for Housing dataset}
  \label{fig:housing_graph_1}
\end{figure}

\begin{table}
  \centering\caption{Discretization result of Housing dataset based on the graph in Figure~\ref{fig:housing_graph_1}. The first twelve rows show the numbers of intervals after discretization by the three method. The thirteenth row is the log-likelihood of the dataset based on the discretizatio results. The heuristic method is over sensitive and discretizes some variables into too many intervals. The MDL method is less sensitive and discretizes most variables by only one interval. The optimal Bayesian approach has a good control on number of intervals. \tim{Fig. vs. Figure}}
  \scalebox{1.0}{
  \input{table_housing_1.tex}
  }

@@

\begin{figure}[ht]
    % \begin{tabular}{cc}
      \input{graph_housing_exp1_distr_3_5}
    % \end{tabular}
  \caption{The discretization result on variable $C$ and $E$ of the network in Figure~\ref{fig:housing_graph_1}. The scatter points are jittered due to the repeated values in the dataset. Furthermore, these repeated values lead to high number of discretization edges, since repeated values can easily form discretized areas by their own.}
  \label{fig:housing_exp1_distr_3_5}
\end{figure}

\begin{figure}[ht]
     %\begin{tabular}{cc}
    \input{graph_housing_exp1_distr_8_5}
     %\end{tabular}
  \caption{The discretization result on variable $H$ and $E$ of the network in Figure~\ref{fig:housing_graph_1}. Data points are not over repeated on $H$, thus the number of edges on $H$ remains desirable.}
  \label{fig:housing_exp1_distr_8_5}
\end{figure}

@@

\begin{figure}[ht]
    % \begin{tabular}{cc}
      \input{graph_housing_2}
    % \end{tabular}
  \caption{The learned discrete-valued Bayesian network from the Housing dataset. Variables $C$, $E$, $J$, $K$ have high numbers of intervals after discretization, since they are in a Markov blanket.
  \tim{Var with 1 interval?}
  }
  \label{fig:housing_graph_2}
\end{figure}

@@

\begin{figure}[ht]
     %\begin{tabular}{cc}
      \input{graph_housing_exp2_distr_3_5}
     %\end{tabular}
  \caption{The discretization result on variable $C$ and $E$ of the network in Figure~\ref{fig:housing_graph_2}. The high number of discretization edges is due to the repeated values on the dataset in variable $C$ and $E$.}
  \label{fig:housing_exp2_distr_3_5}
\end{figure}

@@

To quantatively discuss the sensitivity of the three methods, consider the Bayesian network in Figure \ref{fig:exp_discuss}, where $X$ is continuous-valued and $P_1$ and $P_2$ is discrete-valued. Let $P = \{  P_1, P_2 \}$. Assume $D_X = \{1.0,2.0,3.0, \ldots N \}$. Therefore, the corresponding objective functions to discretize $X$ in the three methods are:
\begin{equation}
\begin{aligned}
\text{Obj}_{\text{MDL}} & = Z_1 \cdot k + \log(N) \cdot \sum^k_{i=1} \sum^{J_P}_{j=1} I(a^*_i, (p_1,p_2)_j) \\
\text{Obj}_{\text{Bayesian}} &= Z_2 \cdot k + \log {{n_i + J_P - 1}\choose{J_P -1}} + \sum^k_{i=1} \log \left(   {{n_i !}\over{n^{(P)}_{i,1}! n^{(P)}_{i,2}! \cdots n^{(P)}_{i,J_P}!    }} \right) \\
\text{Obj}_{\text{Heuristic}} &= Z_2 \cdot k + \log {{n_i + J_{P_1} - 1}\choose{J_{P_1} -1}} + \log {{n_i + J_{P_2} - 1}\choose{J_{P_2} -1}} +\\
& \sum^k_{i=1}  \left[ \log \left(   {{n_i !}\over{n^{(P_1)}_{i,1}! n^{(P_1)}_{i,2}! \cdots n^{(P_1)}_{i,J_{P_1}}    }} \right)  + \log \left(   {{n_i !}\over{n^{(P_2)}_{i,1}! n^{(P_2)}_{i,2} \cdots n^{(P_2)}_{i,J_{P_2}} }} \right) \right]
\end{aligned}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}
